You are the way you structurally talk: Structural-temporal neighborhoods of posts to characterize users in online forums

### 2016-03-06 doc/paper.tex
paper avanzado con la idea bayesiana de detectar cortes de dinamica

\begin{document}
\title{Structural-temporal neighborhoods of posts to characterize users in online forums}

@@

\section{Discussion trees}
A very natural way to represent a conversation thread is a tree graph where vertex represent posts and edges represent replies from some post to another. The tree is rooted at the post who started the thread. This representation allows to apply the mathematical toolbox of graph theory. Figure~\ref{fig:trees} shows some real examples of trees in a Reddit\footnote{www.reddit.com} forum.

@@

\section{Neighborhoods in discussion trees}
In this section, we discuss about the convenience of classic neighbourhood definition in discussion trees. Then we introduce a new definition to overcome some of the limitations of the classic definition when dealing with temporal trees.

\begin{definition}
Given a tree graph $G$, the \textit{structural neighborhood} of distance $d$ of post $i$ is the induced graph composed of all the vertices that are at distance equal or less than $d$ from post $i$
\end{definition}

From this definition a structuarl neighborhood at any distance has an unbounded number of posts. Figure \ref{fig:large_neighboorhood}, for instance, shows a typical case where a post has a large neighborhood even at distance two since a post can only one parent but infinite children (replies).

@@

\begin{figure}
	\centering
	\includegraphics[width=0.4\textwidth]{large_neighborhood}
	\includegraphics[width=0.4\textwidth]{small_neighborhood}
	\caption{The size of the neighborhood at distance $R$ is unbounded. These two graph represent frequent neighborhoods of a post (red) that replied to the root (white). However, the tree in the left corresponds to a very successful root while the three in the right has not brought the attention of too many users.}
	\label{fig:large_neighboorhood}
\end{figure}
\subsection{Structuro-temporal neighborhoods in discussion trees}
In order to reduce the space of possible neighborhoods we introduce the notion of \textit{temporal neighborhood}. Let $t_i$ the time where post $p_i$ was added to the thread $G$. Let $\tau_{ij}=|t_i-t_j|$ be the \textit{temporal distance} between two posts $p_i$ and $p_j$. The \textit{temporal neighborhood} $N_{G}^{(t)}(i, r, n)$of spacial radius $r$ and order $n$ of a post $p_i$ is the set of $n$ posts in its spacial neighborhood $N_{G}(i, r)$ that are temporally closest to $i$ and that can reach $i$ trough posts that are also in the temporal neighborhood.

Finally, we give their own colors the post $i$ and to the root, and we consider that two neighborhoods are equivalent if they are color-isomorphic.

With the above definitions in mind, and for a given forum, we can build a matrix $M$ with the counts of neighborhoods for every user.

\section{Application to Reddit forums}

We analyze the Podemos forum\footnote{https://www.reddit.com/r/podemos} in Reddit. It was was first used as a tool of internal democracy. Today, its members use it also to share and discuss about political news.
The Podemos forum contains 995 discussion threads that started on January. These threads contain a total of 12912 posts written by 1218 users.

\subsection{Counting neighorhoods}
In this section we will build a matrix $M$ for one of the forums of Reddit and then we will cluster users to find groups that have similar preferences over the different types of conversations represented by the neighborhood structures.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{neighborhoods_4_4}
	\caption{Neighborhood counts with $r=4$ and $n=4$}
	\label{fig:neighborhoods_4_4}
\end{figure}

>>>>

\begin{document}
\title{You are the way you structurally talk: Structural-temporal neighborhoods of posts to characterize users in online forums}

@@

\section{Discussion trees}
A natural way to represent a conversation thread is a tree graph where vertex represent posts and edges represent replies from some post to another. The tree is rooted at the post who started the thread. This representation allows to apply the mathematical toolbox of graph theory. Figure~\ref{fig:trees} shows some real examples of trees in a Reddit\footnote{www.reddit.com} forum.

@@

\section{Neighborhoods in discussion trees}

Our goal is to characterize conversations by their local structure. By local structure we mean the structure of the graph that surrounds a given post, which we call \textit{neighbourhood}. Given a set of posts written by a user, analysing the neighbourhoods where user posts are embedded will give us an insight of what kind of user she is. It is important to note that, if the space of possible neighbourhoods is too small, we might not be able to say anything interesting, while if it is too large we might have neighbourhoods that correspond to the same type of conversation.

In this section, we discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees. Then we introduce a new definition to overcome some of the limitations of the classic one when dealing with temporal trees.

@@

\begin{definition}
Given a tree graph $G$, the \textit{structural neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed of all the vertices that are at distance equal or less than $d$ from post $i$.
\end{definition}
%The above definition is an extension of the classic definition of \textit{neighbourhood} which only considers adjacent vertices.

% pros and cons
The above definition has two drawbacks when used in the context of conversation trees.
%%% not all dynamics are captured
First, the dynamics of the conversation (time or order in which posts are attached to the tree) are not entirely captured in the structure of a tree representation. We know, for instance, that the time in which a node was attached to the tree is always posterior to that of its parent. But it impossible to say, by just looking at the structure of the tree, the order in which a set of sibling posts replied to their common parent post. Since these siblings may get more replies themselves, a single \textit{structural neighbourhood} may sometimes correspond to very different dynamics (see Figure).
%TODO: ejemplo artificial para demostrar que la definicion de vecindario estructural puede corresponder a dinamicas muy diferentes.
%% unbounded size
Second, a structural neighbourhood at a given radius $r$ has an unbounded number of posts, and therefore the space of possible neighbourhoods is infinite. This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically equivalent (see Figure)

\begin{figure}
	\centering
	\includegraphics[width=0.4\textwidth]{large_neighborhood}
	\includegraphics[width=0.4\textwidth]{small_neighborhood}
	\caption{The size of the neighbourhood with radius $r$ is unbounded. These two graph represent frequent neighbourhoods of a post (red) that replied to the root (white). However, the tree in the left corresponds to a very successful root while the three in the right has not brought the attention of too many users.}
	\label{fig:large_neighbourhood}
\end{figure}

\subsection{Structuro-temporal neighborhoods in discussion trees}
\begin{definition}
Given a tree graph $G$, the \textit{structural-temporal neighborhood} of distance $d$ of post $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighborhood composed the $n$ vertices that are closer to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.  
\end{definition}

The above definition has two advantages over the previous one. First, dynamics are better taken into account since the neighbourhood only includes posts that are structurally and temporally close. Second, the size of the neighborhood has an upper bound of $max(\mathcal{N}_i(r)), o)$.

Note that this definition does not consider the timestamp of posts but only the order in which they arrived. An alternative definition of structural-temporal neighborhood might include not the $o$  posts that are closest to the ego post, but every post $j$ at distance $|t_i-t_j|<\tau$ from the ego post $i$. However, both options consider that the local dynamic ends at a fixed horizon ($o$ or $\tau$). This is specially problematic for the case of $\tau$ since different conversation threads may grow at different speeds, as well as different parts of the same thread or even the same part of the thread at different moments (e.g. threads barely grow at night). If we were to use $\tau$ as temporal horizon, we need a mechanism to automatically decide the proper $\tau$ for every post.

\begin{definition}
Let $s={p^{(1)},...p^{(m)}}$ be the sequence of posts in the structural neighbourhood $\mathcal{N}_i(r)$ ordered by their temporal distance to $i$. Let $\pi=\{\Delta t_1,...,\Delta t_M\}$ be the sequence of time differences between a post in $s$ and its predecessor in $s$. The \textit{temporal radius} $\tau$ of the neighbourhood is the timestamp of the last post $k$ whose $\Delta t_k$ comes from the same distribution than its predecessors $\Delta t_{k-1},..., \Delta t_{1}$ (an intuition for this definition is given in Figure~\ref{fig:cutpoints})
\end{definition}

According to the above definition, to find the temporal radius we need to find the first outlier in a sequence of $\Delta t$. Assuming the $\Delta t_i$ comes from an Exponential distribution, and that its $\lambda$ parameter comes itself from a Gamma distribution, we can exploit this conjugacy and obtain the analytical form of the probability of a given $\Delta t_k$ of belonging to the same distribution than its predecessors:
\begin{equation}
p(\Delta t_k | \Delta t_1,...\Delta t_{k-1})
=
\int p(\Delta t_k | \lambda_{k-1})p(\lambda_{k-1} |\Delta t_1,...\Delta t_{k-1})
=\frac{(\beta + n\overline{x})^{\alpha+n}}{(\beta + n\overline{x}+\Delta t_k)^{\alpha+n+1}}(\alpha+n)
\end{equation}
Details of this derivation are given in the Appendix.

\begin{figure}
\centering
\includegraphics[width=1\textwidth]{temporal_cutpoint}
\caption{Intuition to choice the temporal cutpoint of a neighbourhood. The two posts out of the box have a distance (time) to their parents clearly bigger than the average, so we consider that they are not part of the same local dynamic.}
\label{fig:cutpoints}
\end{figure}

\subsection{Colored neighborhoods}
In conversation trees, the root vertex plays an important role since it represents the post that opened the conversation and thus set the topic of the thread. Therefore, a reply to the root post is clearly different from a reply to any other post in the tree. We distinguish root posts from non-root posts by assigning root posts a unique color (white). We also assign a unique color to the ego post (red).

\section{Application to Reddit forums}
In this section, we use our definition of structural-temporal neighbourhood to detect different types of users according to the neighbourhoods in which their posts are embedded.

\subsection{Podemos forum}
Our dataset consists of all posts from March 2014 to May 2015 of the Podemos forum in Reddit\footnote{https://www.reddit.com/r/podemos}\footnote{I have also the data for \textit{gameofthrones, france, datascience, machinelearning complexsystems, philosophy, twoxchromosomes, trees, sex}}. The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid the October 18th and 19th 2014. Nowadays, its members use it mainly to share and discuss about political news.
The forum contains 83,6119 posts spread over 47,803 threads and written by 26,193 users. (see Figure~\ref{fig:podemos_distributions})

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{threads_length}%
\includegraphics[width=0.5\textwidth]{users_posts}
\includegraphics[width=0.5\textwidth]{threads_length_cum}%
\includegraphics[width=0.5\textwidth]{users_posts_cum}
\caption{Log-log and cumulative distributions and of thread lengths and posts per user.}
\label{fig:podemos_distributions}
\end{figure}

    \noindent\textcolor{red}{\rule{16cm}{1mm}}
(Il fait que j'implemente l'équation 1 pour chercher les neighborhoods. Puis il faut voir si elle sert tout à fait à bien detecter la baisse du rythme)
\subsection{Counting neighorhoods}
In this section we will build a matrix $M$ for one of the forums of Reddit and then we will cluster users to find groups that have similar preferences over the different types of conversations represented by the neighbourhood structures.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{neighborhoods_4_4}
	\caption{Most frequent neighborhoods with $r=4$ and $n=4$}
	\label{fig:neighborhoods_4_4}
\end{figure}
